The forestecology R package for fitting and assessing neighborhood models of the effect of interspecific competition on the growth of trees

### 2021-09-01 paper/paper.Rmd
Cleaned directory

# knitr settings
opts_chunk$set(
  # Code output:
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  cache = TRUE,
  # Figure:
  out.width = "100%",
  fig.path = "Figures/",
  fig.width = 16 / 2.5,
  fig.height = 9 / 2.5,
  fig.align = "center",
  fig.show = "hold",
  # Etc:
  collapse = TRUE,
  comment = "##",
  tidy.opts = list(width.cutoff = 72),
  tidy = TRUE
)

@@

@@

\begin{equation}
\label{eq:model}
y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot x^{\text{comp}}_{ijk} + \epsilon_{ij}
\end{equation}

@@

\begin{eqnarray}
\label{eq:permutation-hypothesis-test}
&&H_0: \lambda_{jk} = \lambda_{j} \mbox{ for all } k = 1, \ldots, K\\
\text{vs.}&&H_A: \text{at least one } \lambda_{jk} \mbox{ is different} \nonumber
\end{eqnarray}

@@

Our model of interspecific competition relies on a spatial definition of who competitor trees are: all trees within a distance `comp_dist` of a focal tree. Here we set `comp_dist` to 7.5m, a value informed by other studies [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006], but the package could also be used to compare multiple distances and see which is best supported (see Appendix \ref{mult-dist}). We use `comp_dist` and a manually constructed `sf` representation of the study region's boundary as inputs to `add_buffer_variable()` to add a `buffer` boolean variable to `growth_scbi`. All trees with `buffer` equal to `FALSE` will be our focal trees whose growth will be modeled, whereas those with `TRUE` will only act as competitor trees.

@@

The second element of spatial information we add are blocks corresponding to folds of a spatial cross-validation algorithm. Conventional cross-validation algorithms assign individual observations to folds by randomly resampling them all while assuming they are statistically independent. In the case of forest census data however, observations exhibit spatial autocorrelation. We therefore incorporate this dependence into the cross-validation algorithm by resampling spatial blocks of trees [@roberts_cross-validation_2017 @pohjankukka_estimating_2017].

We first manually define an `sf` object defining four folds that partition the study region. We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct `foldID` [@valavi_blockcv_2019]. This `foldID` variable will be used in Section \ref{spatial-cross-validation}.

Figure \ref{fig:scbi-spatial-information} illustrates the net effect of adding these two elements of spatial information to `growth_scbi`.

@@

```{r scbi-spatial-information, out.width="66%", fig.cap="Step 2 - Add spatial information. A buffer region and spatial cross-validation blocks 1 through 4. The location of each tree is marked with its fold number where the folds are delineated with solid lines. The color of each digit indicates whether the tree is part of the buffer region (thus will only be considered as a competitor tree) or is part of the interior of the study region (thus is a focal tree whose growth is of modeled interest)."}
ggplot() +
  geom_sf(data = blocks_scbi, fill = "transparent", linetype = "dashed") +
  geom_sf_text(
    data = growth_scbi %>% sample_n(1000),
    aes(label = foldID, col = buffer)
  )
```

@@

We then identify all focal trees and their corresponding competitor trees. More specifically, identify all trees that are not part of the buffer region, have a valid `growth` measurement, and have at least one neighbor within `r comp_dist`m. We do this using `create_focal_vs_comp()`, which takes the previously detailed `comp_dist` and `id` arguments, the `sf` representation of the spatial cross-validation blocks `blocks_scbi`, and a specification `comp_x_var` of the `basal_area` variable we use as the competitor explanatory variable $x^{\text{comp}}_{ijk}$ from Equation \ref{eq:model}. This function returns a new data frame `focal_vs_comp_scbi`.

```{r scbi-focal-vs-comp}
focal_vs_comp_scbi <- growth_scbi %>%
  create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID", comp_x_var = "basal_area")
focal_vs_comp_scbi %>%
  select(focal_ID, focal_sp, geometry, growth, comp)
```

@@

```{r scbi-model-predict}
focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
  mutate(growth_hat = predict(comp_bayes_lm_scbi,
    newdata = focal_vs_comp_scbi
  ))
```
```{r show-focal-vs-comp}
focal_vs_comp_scbi %>%
  select(focal_ID, focal_sp, dbh, growth, growth_hat)
```

@@

plot1 <- autoplot(comp_bayes_lm_scbi,
  type = "intercepts",
  sp_to_plot = sp_to_plot
)
plot2 <- autoplot(comp_bayes_lm_scbi,
  type = "dbh_slopes",
  sp_to_plot = sp_to_plot
)
plot3 <- autoplot(comp_bayes_lm_scbi,
  type = "competition",
  sp_to_plot = sp_to_plot
)

@@

focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
  mutate(growth_hat_shuffle = predict(comp_bayes_lm_scbi_shuffle,
    newdata = focal_vs_comp_scbi
  ))
```
```{r show-rmse-shuffle}
model_rmse_shuffle <- focal_vs_comp_scbi %>%
  rmse(truth = growth, estimate = growth_hat_shuffle) %>%
  pull(.estimate)
model_rmse_shuffle
```

@@

`run_cv()` also accepts the `run_shuffle` argument in order to permute competitor species identity as described in Section \ref{permutation-test}. Figure \ref{fig:scbi-simulation} compares model performance for 49 permutations of competitor species and RMSE calculations, both with and without cross-validation. Without cross-validation, competitor species identity does matter as the observed RMSE was significantly lower than the permutation null distribution of RMSE. However, once we incorporate spatial cross-validation, this improvement disappears. These results suggest that in this 9 ha subplot of the SCBI plot, competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other locations [@allen_permutation_2020 @uriarte_spatially_2004]. This provides a striking example of the importance of cross-validation, as without it the over-fit model gives rise to an incorrect conclusion.

@@

We also hope to extend the `forestecology` package's functionality to account for a larger variety of models for tree growth. One clear future direction would be to allow competition based on species trait values rather than species identity. There is evidence that traits predict competitive outcomes [@kunstler_competitive_2012, @lasky_trait-mediated_2014, @uriarte_trait_2010]. Thus an extension of the model would allow $\lambda$ values from Equation \ref{eq:model} to be a function of the traits of competing species.

Lastly, the `forestecology` current uses the `blockCV` package behind the scenes to create the spatial blocks acting as folds for our spatial cross-validation algorithm detailed in Sections \ref{spatial-information} and \ref{spatial-cross-validation} [@valavi_blockcv_2019]. This back-end functionality could be substituted with the `spatialsample` package for spatial resampling infrastructure; a `tidymodels` package under active development as of 2021 [@spatialsample_package, @tidymodels_package].

Lastly, currently the package only implements the Bayesian linear regression model detailed in Equation \ref{eq:model}. As we demonstrate in Section \ref{model-fit-predict} however, the fitting of this model is self-contained in a single function `comp_bayes_lm()` which returns an object of S3 class type `comp_bayes_lm`. This class has generic methods implemented to print, make predictions, and plot all results. Therefore the package can be modularly extended to fit other models as long as they are coded similarly to `comp_bayes_lm()` and have equivalent generic methods implemented.

# Acknowledgments

We thank Sophie Li for their feedback on the package interface. The authors declare no conflicts of interest.

# Author's contributions

AYK and DNA conceived the ideas and coded a draft of the package. AYK wrote an initial manuscript draft. SPC rewrote much of the package's code to align with R and "tidy" best practices [@wickham_welcome_2019]. All authors contributed to subsequent drafts and gave final approval for manuscript.

# Data accessibility

We intend to archive all data and source code for this manuscript on GitHub at https://github.com/rudeboybert/forestecology and on Zenodo upon acceptance. The example Smithsonian Conservation Biology Institute census and species information data are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/ and are archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].

@@

\appendix

@@

For all the above analyses we set the cut off distance (`comp_dist`) for two stems to compete as `r comp_dist`m. This distance has been estimated in previous neighborhood competition studies in forests [@canham_neighborhood_2004, @uriarte_spatially_2004, @canham_neighborhood_2006]. We used `r comp_dist`m in @allen_permutation_2020 as an average of the values estimated in other studies. But our package can be used to find which distance is best supported by the data. Here we provide an example using another section of the SCBI plot to provide an additional example of the cross-validation block layout. To speed computation we do not consider species differences in competitive effects and treat all species as the same.

We observe in Figure \ref{fig:appendix2-plot} that a cut-off distance of approximately 6m minimizes the cross-validation estimated RMSE.

@@

growth_scbi <-
  compute_growth(
    census_1 = census_2013_scbi,
    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
    id = "stemID"
  ) %>%
  # make all species the same, needs to be factor with at least two levels
  mutate(sp = factor("A", levels = c("A", "B"))) %>%
  # Compute basal area:
  mutate(basal_area = 0.0001 * pi * (dbh1 / 2)^2)

@@

  focal_vs_comp_scbi <- growth_scbi %>%
    create_focal_vs_comp(comp_dist = comp_dist, blocks = blocks_scbi, id = "stemID", comp_x_var = "basal_area") %>%
    run_cv(comp_dist = comp_dist, blocks = blocks_scbi)

@@

# Compute growth of trees based on census data -------------------------
census_2013_scbi <- here("paper/scbi.stem2.csv") %>%
  read_csv() %>%
  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
  mutate(
    date = mdy(date),
    dbh = as.numeric(dbh) / 10
  ) %>%
  filter(gx < 300, between(gy, 300, 600))

@@

# Add spatial information ----------------------------------------------
# Define buffer region using competitive distance range
comp_dist <- 7.5

@@

# Compute focal versus competitor tree information ---------------------
focal_vs_comp_scbi <- growth_scbi %>%
  create_focal_vs_comp(comp_dist, blocks = blocks_scbi, id = "stemID", comp_x_var = "basal_area")

# Fit model and make predictions ---------------------------------------
# Number of permutation shuffles:
num_shuffle <- 49

@@

# 1. Compute observed test statistic: RMSE with no cross-validation ----
# Fit model (compute posterior parameters)
comp_bayes_lm_scbi <- focal_vs_comp_scbi %>%
  comp_bayes_lm(prior_param = NULL, run_shuffle = FALSE)

@@

# 2. Compute observed test statistic: RMSE with cross-validation -------
observed_RMSE_CV <- focal_vs_comp_scbi %>%
  run_cv(comp_dist = comp_dist, blocks = blocks_scbi) %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)

# 3. Permutation distribution: RMSE with no cross-validation -----------
# Compute num_shuffle permutation test statistics
shuffle_RMSE <- numeric(length = num_shuffle)
for (j in 1:num_shuffle) {
  # Fit model (compute posterior parameters) with shuffling
  comp_bayes_lm_scbi <- focal_vs_comp_scbi %>%
    comp_bayes_lm(prior_param = NULL, run_shuffle = TRUE)

@@

# 4. Permutation distribution: RMSE with cross-validation --------------
# Compute num_shuffle permutation test statistics
shuffle_RMSE_CV <- numeric(length = num_shuffle)

@@

# Visualize results ----------------------------------------------------
model_comp <- bind_rows(
  model_comp_tbl %>%
    select(run_time,
      observed = observed_RMSE,
      shuffle = shuffle_RMSE
    ) %>%
    mutate(CV = FALSE),
  model_comp_tbl %>%
    select(run_time,
      observed = observed_RMSE_CV,
      shuffle = shuffle_RMSE_CV
    ) %>%
    mutate(CV = TRUE)
) %>%
  gather(type, RMSE, -c(run_time, CV))

@@

cv_plot <- ggplot() +
  geom_vline(
    data = model_comp_observed,
    aes(xintercept = RMSE, col = CV),
    linetype = "dashed", show.legend = F
  ) +
  geom_histogram(
    data = model_comp_shuffle,
    aes(x = RMSE, fill = CV), bins = 50
  ) +
  labs(
    fill = "Cross-validated?",
    x = expression(paste("RMSE (cm ", y^{
      -1
    }, ")"))
  ) +
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_fill_viridis(discrete = TRUE) +
  theme_light()
cv_plot

>>>>

# knitr settings
opts_chunk$set(
  # Code output:
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  cache = TRUE,
  # Figure:
  out.width = "100%",
  fig.path = "figures/",
  fig.width = 16 / 2.5,
  fig.height = 9 / 2.5,
  fig.align = "center",
  fig.show = "hold",
  # Etc:
  collapse = TRUE,
  comment = "##",
  tidy.opts = list(width.cutoff = 72),
  tidy = FALSE
)

@@

# Abstract {-}

1. Neighborhood competition models are powerful tools to measure the effect of interspecific competition. Statistical methods to ease the application of these models are currently lacking.
2. We present the `forestecology` package providing methods to i) specify neighborhood competition models, ii) evaluate the effect of competitor species identity using permutation tests, and iii) measure model performance using spatial cross-validation.  Following @allen_permutation_2020, we implement a Bayesian linear regression neighborhood competition model.
3. We demonstrate the package's functionality using data from the Smithsonian Conservation Biology Institute's large forest dynamics plot, part of the ForestGEO global network of research sites. Given ForestGEO's data collection protocols and data formatting standards, the package was designed with cross-site compatibility in mind. We highlight the importance of spatial cross-validation when interpreting model results. 
4. The package features i) `tidyverse`-like structure whereby verb-named functions can be modularly "piped" in sequence, ii) functions with standardized inputs/outputs of simple features `sf` package class, and iii) an S3 object-oriented implementation of the Bayesian linear regression model. These three facts allow for clear articulation of all the steps in the sequence of analysis and easy wrangling and visualization of the geospatial data. Furthermore, while the package only has Bayesian linear regression implemented, the package was designed with extensibility to other methods in mind.

Keywords: forest ecology, interspecific competition, neighborhood competition, tree growth, R, ForestGEO, spatial cross-validation

@@

\begin{eqnarray}
y_{ij} = \beta_{0,j} + \beta_{\text{dbh},j} \cdot \text{dbh}_{ij} + \sum_{k=1}^{K} \lambda_{jk} \cdot x^{\text{comp}}_{ijk} + \epsilon_{ij} (\#eq:model)
\end{eqnarray}

@@

\begin{eqnarray}
&&H_0: \lambda_{jk} = \lambda_{j} \mbox{ for all } k = 1, \ldots, K (\#eq:permutation-hypothesis-test)\\
\text{vs.}&&H_A: \text{at least one } \lambda_{jk} \mbox{ is different} \nonumber
\end{eqnarray}

@@

Our model of interspecific competition relies on a spatial definition of who competitor trees are: all trees within a distance `comp_dist` of a focal tree. Here we set `comp_dist` to 7.5m, a value informed by other studies [@canham_neighborhood_2004; @uriarte_spatially_2004; @canham_neighborhood_2006], but the package could also be used to compare multiple distances and see which is best supported (see Appendix \@ref(mult-dist)). We use `comp_dist` and a manually constructed `sf` representation of the study region's boundary as inputs to `add_buffer_variable()` to add a `buffer` boolean variable to `growth_scbi`. All trees with `buffer` equal to `FALSE` will be our focal trees whose growth will be modeled, whereas those with `TRUE` will only act as competitor trees.

@@

The second element of spatial information we add are blocks corresponding to folds of a spatial cross-validation algorithm. Conventional cross-validation algorithms assign individual observations to folds by randomly resampling them all while assuming they are statistically independent. In the case of forest census data however, observations exhibit spatial autocorrelation. We therefore incorporate this dependence into the cross-validation algorithm by resampling spatial blocks of trees [@roberts_cross-validation_2017; @pohjankukka_estimating_2017].

We first manually define an `sf` object defining four folds that partition the study region. We then use the output of the `spatialBlock()` function from the `blockCV` package to associate each tree in `growth_scbi` to the correct `foldID` [@valavi_blockcv_2019]. This `foldID` variable will be used in Section \@ref(spatial-cross-validation).

Figure \@ref(fig:scbi-spatial-information) illustrates the net effect of adding these two elements of spatial information to `growth_scbi`.

@@

```{r scbi-spatial-information, out.width="66%", fig.cap="Step 2 - Add spatial information. A buffer region and spatial cross-validation blocks 1 through 4. The location of each tree is marked with its fold number where the folds are delineated with solid lines. The color of each digit indicates whether the tree is part of the buffer region (thus will only be considered as a competitor tree) or is part of the interior of the study region (thus is a focal tree whose growth is of modeled interest)."}
ggplot() +
  geom_sf(
    data = blocks_scbi, 
    fill = "transparent", linetype = "dashed"
  ) +
  geom_sf_text(
    data = growth_scbi %>% sample_n(1000),
    aes(label = foldID, col = buffer)
  )
```

@@

We then identify all focal trees and their corresponding competitor trees. More specifically, identify all trees that are not part of the buffer region, have a valid `growth` measurement, and have at least one neighbor within `r comp_dist`m. We do this using `create_focal_vs_comp()`, which takes the previously detailed `comp_dist` and `id` arguments, the `sf` representation of the spatial cross-validation blocks `blocks_scbi`, and a specification `comp_x_var` of the `basal_area` variable we use as the competitor explanatory variable $x^{\text{comp}}_{ijk}$ from Equation \@ref(eq:model). This function returns a new data frame `focal_vs_comp_scbi`.

```{r scbi-focal-vs-comp}
focal_vs_comp_scbi <- growth_scbi %>%
  create_focal_vs_comp(
    comp_dist, blocks = blocks_scbi, id = "stemID", 
    comp_x_var = "basal_area"
  )

focal_vs_comp_scbi %>%
  select(focal_ID, focal_sp, geometry, growth, comp)
```

@@

```{r scbi-model-predict}
focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
  mutate(
    growth_hat = predict(comp_bayes_lm_scbi,
                         newdata = focal_vs_comp_scbi)
    )
```
```{r show-focal-vs-comp}
focal_vs_comp_scbi %>%
  select(focal_ID, focal_sp, dbh, growth, growth_hat)
```

@@

plot1 <- autoplot(
  comp_bayes_lm_scbi,
  type = "intercepts",
  sp_to_plot = sp_to_plot
)
plot2 <- autoplot(
  comp_bayes_lm_scbi,
  type = "dbh_slopes",
  sp_to_plot = sp_to_plot
)
plot3 <- autoplot(
  comp_bayes_lm_scbi,
  type = "competition",
  sp_to_plot = sp_to_plot
)

@@

focal_vs_comp_scbi <- focal_vs_comp_scbi %>%
  mutate(
    growth_hat_shuffle = predict(comp_bayes_lm_scbi_shuffle,
                                 newdata = focal_vs_comp_scbi)
  )
```
```{r show-rmse-shuffle}
model_rmse_shuffle <- focal_vs_comp_scbi %>%
  rmse(truth = growth, estimate = growth_hat_shuffle) %>%
  pull(.estimate)
model_rmse_shuffle
```

@@

`run_cv()` also accepts the `run_shuffle` argument in order to permute competitor species identity as described in Section \@ref(permutation-test). Figure \@ref(fig:scbi-simulation) compares model performance for 49 permutations of competitor species and RMSE calculations, both with and without cross-validation. Without cross-validation, competitor species identity does matter as the observed RMSE was significantly lower than the permutation null distribution of RMSE. However, once we incorporate spatial cross-validation, this improvement disappears. These results suggest that in this 9 ha subplot of the SCBI plot, competitive interactions do not depend on the identity of the competitor, which is the opposite of what has been observed in other locations [@allen_permutation_2020; @uriarte_spatially_2004]. This provides a striking example of the importance of cross-validation, as without it the over-fit model gives rise to an incorrect conclusion.

@@

We also hope to extend the `forestecology` package's functionality to account for a larger variety of models for tree growth. One clear future direction would be to allow competition based on species trait values rather than species identity. There is evidence that traits predict competitive outcomes [@kunstler_competitive_2012; @lasky_trait-mediated_2014; @uriarte_trait_2010]. Thus an extension of the model would allow $\lambda$ values from Equation \@ref(eq:model) to be a function of the traits of competing species.

Lastly, the `forestecology` current uses the `blockCV` package behind the scenes to create the spatial blocks acting as folds for our spatial cross-validation algorithm detailed in Sections \@ref(spatial-information) and \@ref(spatial-cross-validation) [@valavi_blockcv_2019]. This back-end functionality could be substituted with the `spatialsample` package for spatial resampling infrastructure; a `tidymodels` package under active development as of 2021 [@spatialsample_package; @tidymodels_package].

Lastly, currently the package only implements the Bayesian linear regression model detailed in Equation \@ref(eq:model). As we demonstrate in Section \@ref(model-fit-predict) however, the fitting of this model is self-contained in a single function `comp_bayes_lm()` which returns an object of S3 class type `comp_bayes_lm`. This class has generic methods implemented to print, make predictions, and plot all results. Therefore the package can be modularly extended to fit other models as long as they are coded similarly to `comp_bayes_lm()` and have equivalent generic methods implemented.

# Acknowledgements {-}

We thank Sophie Li for their feedback on the package interface.

# Conflict of Interest {-}

We declare no competing interests.

# Author's contributions {-}

**Albert Y. Kim**: Conceptualization (equal); Methodology (equal); Software (equal); Writing-original draft (lead); Writing-review & editing (equal). **David N. Allen**: Conceptualization (equal); Methodology (equal); Software (equal); Writing-original draft (supporting); Writing-review & editing (equal). **Simon P. Couch**: Software (supporting); Writing-review & editing (supporting).

# Data accessibility {-}

All data and source code for this manuscript on GitHub at https://github.com/rudeboybert/forestecology and are archived on Zenodo at https://doi.org/10.5281/zenodo.5348060. The example Smithsonian Conservation Biology Institute census and species information data are available on GitHub at https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/tree/master/ and are archived on Zenodo at https://doi.org/10.5281/zenodo.2649301 [@gonzalez-akre_scbi-forestgeoscbi-forestgeo-data_2020].

@@

@@

For all the above analyses we set the cut off distance (`comp_dist`) for two stems to compete as `r comp_dist`m. This distance has been estimated in previous neighborhood competition studies in forests [@canham_neighborhood_2004; @uriarte_spatially_2004; @canham_neighborhood_2006]. We used `r comp_dist`m in @allen_permutation_2020 as an average of the values estimated in other studies. But our package can be used to find which distance is best supported by the data. Here we provide an example using another section of the SCBI plot to provide an additional example of the cross-validation block layout. To speed computation we do not consider species differences in competitive effects and treat all species as the same.

We observe in Figure \@ref(fig:appendix2-plot) that a cut-off distance of approximately 6m minimizes the cross-validation estimated RMSE.

@@

growth_scbi <-
  compute_growth(
    census_1 = census_2013_scbi,
    census_2 = census_2018_scbi %>% filter(!str_detect(codes, "R")),
    id = "stemID"
  ) %>%
  # make all species the same, needs to be factor with at least 
  # two levels
  mutate(sp = factor("A", levels = c("A", "B"))) %>%
  # Compute basal area:
  mutate(basal_area = 0.0001 * pi * (dbh1 / 2)^2)

@@

  focal_vs_comp_scbi <- growth_scbi %>%
    create_focal_vs_comp(
      comp_dist = comp_dist, blocks = blocks_scbi, id = "stemID", 
      comp_x_var = "basal_area"
    ) %>%
    run_cv(comp_dist = comp_dist, blocks = blocks_scbi)

@@

# Compute growth of trees based on census data ------------------------
census_2013_scbi <- here("paper/scbi.stem2.csv") %>%
  read_csv() %>%
  select(stemID, sp, date = ExactDate, gx, gy, dbh, codes, status) %>%
  mutate(
    date = mdy(date),
    dbh = as.numeric(dbh) / 10
  ) %>%
  filter(gx < 300, between(gy, 300, 600))

@@

# Add spatial information ---------------------------------------------
# Define buffer region using competitive distance range
comp_dist <- 7.5

@@

# Compute focal versus competitor tree information --------------------
focal_vs_comp_scbi <- growth_scbi %>%
  create_focal_vs_comp(
    comp_dist, blocks = blocks_scbi, id = "stemID", 
    comp_x_var = "basal_area"
  )

# Fit model and make predictions --------------------------------------
# Number of permutation shuffles:
num_shuffle <- 49

@@

# 1. Compute observed test statistic: RMSE with no cross-validation ---
# Fit model (compute posterior parameters)
comp_bayes_lm_scbi <- focal_vs_comp_scbi %>%
  comp_bayes_lm(prior_param = NULL, run_shuffle = FALSE)

@@

# 2. Compute observed test statistic: RMSE with cross-validation ------
observed_RMSE_CV <- focal_vs_comp_scbi %>%
  run_cv(comp_dist = comp_dist, blocks = blocks_scbi) %>%
  rmse(truth = growth, estimate = growth_hat) %>%
  pull(.estimate)

# 3. Permutation distribution: RMSE with no cross-validation ----------
# Compute num_shuffle permutation test statistics
shuffle_RMSE <- numeric(length = num_shuffle)
for (j in 1:num_shuffle) {
  # Fit model (compute posterior parameters) with shuffling
  comp_bayes_lm_scbi <- focal_vs_comp_scbi %>%
    comp_bayes_lm(prior_param = NULL, run_shuffle = TRUE)

@@

# 4. Permutation distribution: RMSE with cross-validation -------------
# Compute num_shuffle permutation test statistics
shuffle_RMSE_CV <- numeric(length = num_shuffle)

@@

# Visualize results ---------------------------------------------------
model_comp <- bind_rows(
  model_comp_tbl %>%
    select(run_time,
           observed = observed_RMSE,
           shuffle = shuffle_RMSE
    ) %>%
    mutate(CV = FALSE),
  model_comp_tbl %>%
    select(run_time,
           observed = observed_RMSE_CV,
           shuffle = shuffle_RMSE_CV
    ) %>%
    mutate(CV = TRUE)
) %>%
  gather(type, RMSE, -c(run_time, CV))

@@

cv_plot <- ggplot() +
  geom_vline(
    data = model_comp_observed,
    aes(xintercept = RMSE, col = CV),
    linetype = "dashed", show.legend = F
  ) +
  geom_histogram(
    data = model_comp_shuffle,
    aes(x = RMSE, fill = CV), bins = 50
  ) +
  labs(
    fill = "Cross-validated?",
    x = expression(paste("RMSE (cm ", y^{-1}, ")"))
  ) +
  scale_color_viridis(discrete = TRUE, option = "D") +
  scale_fill_viridis(discrete = TRUE) +
  theme_light()
cv_plot
